Naturalness in U(1) extensions of the MSSM

### 2014-11-11 paper/e6ssm-tuning-paper.tex
added some detail to the model section.

@@

\section{\label{sec:model}U(1) extensions and the E$_6$SSM}
% Table of U(1)' charges, which may or may not be useful to include.
\begin{tabular}{cccccccccccccc}
 & $\hat{Q}$ & $\hat{u}^C$ & $\hat{d}^C$ & $\hat{L}$ & $\hat{e}^C$ & $\hat{N}^C$ & $\hat{S}$ & $\hat{H}_2$ & $\hat{H}_1$ & $\hat{D}$ & $\hat{\overline{D}}$ & $\hat{H}'$ & $\hat{\overline{H'}}$ \\[1mm]
$\sqrt{\frac{5}{3}}Q_i^Y$ & $\frac{1}{6}$ & $-\frac{2}{3}$ & $\frac{1}{3}$ & $-\frac{1}{2}$ & $1$ & $0$ & $0$ & $\frac{1}{2}$ & $-\frac{1}{2}$ & $-\frac{1}{3}$ & $\frac{1}{3}$ & $-\frac{1}{2}$ & $\frac{1}{2}$ \\[1mm]
%\hline
$2\sqrt{6}Q_i^\psi$ & $1$ & $1$ & $1$ & $1$ & $1$ & $1$ & $4$ & $-2$ & $-2$ & $-2$ & $-2$ & $1$ & $-1$\\[1mm]
%\hline
$2\sqrt{10}Q_i^\chi$ & $-1$ & $-1$ & $3$ & $3$ & $-1$ & $-5$ & $0$ & $2$ & $-2$ & $2$ & $-2$ & $3$ & $-3$\\[1mm]
%\hline
$\sqrt{40}Q_i^N$ & $1$ & $1$ & $2$ & $2$ & $1$ & $0$ & $5$ & $-2$ & $-3$ & $-2$ & $-3$ & $2$ & $-2$ \\[1mm]
\caption{The $U(1)_Y$, $U(1)_\psi$, $U(1)_\chi$ and $U(1)_N$ charges of the chiral superfields in the $E_6$ model. The specific case of $U(1)_N$, corresponding to the E$_6$SSM, is obtained for $\theta_{E_6}=\arctan\sqrt{15}$.}
\label{tab:E6charges}

>>>>

% macros
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\ba}{\begin{eqnarray}}
\newcommand{\ea}{\end{eqnarray}}
\newcommand{\SuperField}[1]{\hat{#1}}

@@

\section{\label{sec:model}U(1) extensions and the E$_6$SSM}
In this paper we consider U(1) extensions of the MSSM where the gauge group at low energies is,
 \be
SU(3)_C\times SU(2)_W\times U(1)_Y\times U(1)^\prime, \ee 
where $U(1)^\prime$ is the new gauge group beyond that of the SM and MSSM. The minimal superfield content of $U(1)$ extensions which solve the $\mu$-problem should be ordinary left handed quark, $\hat{Q}_i$ and lepton $\hat{L}_i$ ($i=1,2,3$)superfields along with right handed superfields $\hat{u}^c_i$, $\hat{d}^c_i$,  $\hat{e}^c_i$ ($i=1,2,3$) for the up-type (s)quarks, down-type (s)quarks and charged (s)leptons respectively and three Higgs superfields, up-type $\hat{H}_u$, down-type $\hat{H}_d$ and a singlet under the SM gauge group $\hat{S}$.

  The couplings for the $U(1)^\prime$ gauge group should allow the following renormalisable superpotential terms required in the USSM,
\be
W_{USSM} =  y^U_{ijk} \hat{u}^c_i \hat{H}^u_{j} \hat{Q}_k + y^D_{ijk} \hat{d}^c_i \hat{H}^d_{j} \hat{Q}_k + y^E_{ijk} \hat{e}^c_i \hat{H}^d_{j} \hat{L}_k + \lambda \hat{S} \hat{H}^d \hat{H},
\ee, with $i,j,k \in \{1,2,3\}$.

The $U(1)^\prime$ charges should allow for cancellations of gauge
anomalies. The most elegant way to do this is to use an extra $U(1)$
gauge symmetry that can be obtained from the break down of the $E_6$
gauge symmetry which is anomaly free and have all matter multiplets
from the $Z7$-plet representation of $E_6$ survive down to low energies.

The breaking of $E_6$ into $SO(10)$, gives rise to $E_6\to
SO(10)\times U(1)_{\psi}$, and the subsequent breaking of $SO(10)$
into $SU(5)$, gives $SO(10)\to SU(5)\times U(1)_{\chi}$ (this is
reviewed in e.g.~\cite{Langacker:2008yv})). The extar $U(1)$ gauge symmetry at low energies should then be a linear combination of these in the $E_6$ inspired case,
\be
U(1)^\prime = \cos(\theta) U(1)_{\chi} + \sin(\theta) U(1)_{\psi}.
\ee In table \ref{tab:E6charges} the charges for several popular $E_6$ inspired $U(1)$ extensions are shown.
% Table of U(1)' charges, which may or may not be useful to include.
\begin{table}[h]
\centering
\begin{ruledtabular}
\begin{tabular}{cccccccccccccc}
 & $\hat{Q}$ & $\hat{u}^C$ & $\hat{d}^C$ & $\hat{L}$ & $\hat{e}^C$ & $\hat{N}^C$ & $\hat{S}$ & $\hat{H}_2$ & $\hat{H}_1$ & $\hat{D}$ & $\hat{\overline{D}}$ & $\hat{H}'$ & $\hat{\overline{H'}}$ \\[1mm]
\hline
$\sqrt{\frac{5}{3}}Q_i^Y$ & $\frac{1}{6}$ & $-\frac{2}{3}$ & $\frac{1}{3}$ & $-\frac{1}{2}$ & $1$ & $0$ & $0$ & $\frac{1}{2}$ & $-\frac{1}{2}$ & $-\frac{1}{3}$ & $\frac{1}{3}$ & $-\frac{1}{2}$ & $\frac{1}{2}$ \\[1mm]
%\hline
$2\sqrt{6}Q_i^\psi$ & $1$ & $1$ & $1$ & $1$ & $1$ & $1$ & $4$ & $-2$ & $-2$ & $-2$ & $-2$ & $1$ & $-1$\\[1mm]
%\hline
$2\sqrt{10}Q_i^\chi$ & $-1$ & $-1$ & $3$ & $3$ & $-1$ & $-5$ & $0$ & $2$ & $-2$ & $2$ & $-2$ & $3$ & $-3$\\[1mm]
%\hline
$\sqrt{40}Q_i^N$ & $1$ & $1$ & $2$ & $2$ & $1$ & $0$ & $5$ & $-2$ & $-3$ & $-2$ & $-3$ & $2$ & $-2$ \\[1mm]
\end{tabular}
\end{ruledtabular}
\caption{The $U(1)_Y$, $U(1)_\psi$, $U(1)_\chi$ and $U(1)_N$ charges of the chiral superfields in the $E_6$ model. The specific case of $U(1)_N$, corresponding to the E$_6$SSM, is obtained for $\theta_{E_6}=\arctan\sqrt{15}$.}
\label{tab:E6charges}
\end{table}

To ensure anomaly cancellation in these cases the full matter content
from three generations of $27$-plets should survive to low energies.
the low energy matter content would then be extended to, \be
(\hat{Q}_i,\,\hat{u}^c_i,\,\hat{d}^c_i,\,\hat{L}_i,\,\hat{e}^c_i)
+(\hat{D}_i,\,\hat{\overline{D}}_i)+(\hat{S}_{i})+(\hat{H}^u_i)+(\hat{H}^d_i),\ee
where the $\hat{S}_{i}$, $(\hat{H}^u_i$ and $\hat{H}^d_i$ are three
generations of doublets, whose quantum numbers will be like the
singlet, and up-, down-type Higgs fields and the $\hat{D}_i$ and
$\hat{\overline{D}}_i$ being $SU(3)_C$ triplets that reside in the same $SU(5)$
multiplets as these Higgs-like states.

If one wishes to maintain gauge coupling unification this set of
states should be augmented by two extra $SU(2)$ doublet states
$H^\prime$ and $\overline{H}^\prime$ belonging to other $27^\prime$
and $\overline{27}^\prime$ multiplets that must be incomplete at low
energies.

The full superpotential for $E_6$ inspired models will then be,

\be
W_{E6} = W_0 + W_1 + W_2,
\label{Eq:e6-superpot}
\ee where
\begin{eqnarray}
W_0 &=& \lambda_{ijk} \hat{S}_i \hat{H}^d_{j} \hat{H}^u_{k} + \kappa_{ijk} \hat{S}_i \hat{D}_j \hat{\bar{D}}_k + h^N_{ijk} \hat{N}^c_i \hat{H}^u_{j} \hat{L}_k \nonumber \\
& & + h^U_{ijk} \hat{u}^c_i \hat{H}^u_{j} \hat{Q}_k + h^D_{ijk} \hat{d}^c_i \hat{H}^d_{j} \hat{Q}_k + h^E_{ijk} \hat{e}^c_i \hat{H}^d_{j} \hat{L}_k, \\
W_1 &=& g^Q_{ijk} \hat{D}_i \hat{Q}_j \hat{Q}_k + g^q_{ijk} \hat{\bar{D}}_i \hat{d}^c_j \hat{u}^c_k, \\
W_2 &=& g^N_{ijk} \hat{N}^c_i \hat{D}_j \hat{d}^c_k + g^E_{ijk} \hat{e}^c_i \hat{D}_j \hat{u}^c_k + g^D_{ijk} \hat{Q}_i \hat{L}_j \hat{\bar{D}}_k.
\label{Eq:e6-superpot-parts}
\end{eqnarray}

Nonetheless while this model is very elegant so far, the  superpotential, \ref{Eq:e6-superpot} contains dangerous terms which can induce proton decay and lead to large flavour changing neutral currents (FCNCs).  There are a number of approaches to suppress these terms, involving the use of different discrete symmetries.  Here for the purposes of renormalisation group running we will simply include the following unsuppressed superpotential terms, which follows the approach taking in work on the cE$_6$SSM \cite{Athron:2009ue, Athron:2009bs},
%
%% \begin{align}
%%   \begin{split}
    \ba
    W &\approx& - y_{\tau} (\SuperField{H}_d
    \SuperField{L}_3) \SuperField{e}^C_3 - y_b (\SuperField{H}_d
    \SuperField{Q}_3) \SuperField{d}_3^C - y_t (\SuperField{Q}_3
    \SuperField{H}_u) \SuperField{u}_3^C\\
    &&
    + \lambda_i \SuperField{S} (\SuperField{H}_{1i}
    \SuperField{H}_{2i}) + \kappa_i \SuperField{S} \SuperField{D}_i
    \SuperField{\overline{D}}_i + \mu' (\SuperField{H}'
    \SuperField{\overline{H'}}),
    \ea
  %% \end{split}
  \label{SuPot_RGE}
%%\end{align}
%
